Inferring probabilistic stellar rotation periods using Gaussian processes

### 2017-02-13 documents/ms.tex

The flexibility of this GP model allows for posterior multimodality and
``over-fitting''--like behavior.
For example, if $l$ is small, the non-periodic factor in the covariance
kernel may dominate, allowing for a good fit to the data without
requiring any periodic covariance structure---even if clear periodic
structure is present.
Additionally, if $\Gamma$ is too large, the GP model becomes extremely
flexible and can fit the data without varying the period.
Managing this flexibility to reliably retrieve the period parameter requires
imposing informative priors on the non-period GP parameters.
% While there are no strictly {\it a priori} reasons to believe
% that any of the hyperparameters should take values in any specific range,
% we have found in this work that reasonable priors for any particular
% dataset can be determined by experimentation.
% We note this is a somewhat subjective art, with a goal of
% not allowing the GP to become so flexible that it does not require
% periodicity to fit the data even in the presence of clear periodicity.
In particular, we find it necessary to avoid large values of
$A$ and $\Gamma$, and small values of $l$; though the exact details of what
works well may differ among datasets (see \ref{sec:} and
\ref{sec:} for two different examples of hyperparameter priors).
More physically motivated priors are discussed in \textsection
\ref{sec:discussion}.

@@

We do not attempt to decide which single
peak in the ACF best represents the true rotation period,
but rather we identify several \emph{candidate} periods and define
a weighting scheme in order to create a noncommital, though useful,
multimodal prior.  While this procedure does not avoid heuristic choices,
the fact that its goal is simply to create a \emph{prior} for a
probabilistically justified inference scheme rather than
to identify a single correct period softens the potential impact
of these choices.

>>>>

The flexibility of this GP model allows for posterior multimodality and
``over-fitting''--like behavior.
For example, if $l$ is small, the non-periodic factor in the covariance
kernel may dominate, allowing for a good fit to the data without
requiring any periodic covariance structure---even if clear periodic
structure is present.
Additionally, for large values of $\Gamma$ the GP model becomes extremely
flexible and can fit the data without varying the period.
Managing this flexibility to reliably retrieve the period parameter requires
imposing informative priors on the non-period GP parameters.
% While there are no strictly {\it a priori} reasons to believe
% that any of the hyperparameters should take values in any specific range,
% we have found in this work that reasonable priors for any particular
% dataset can be determined by experimentation.
% We note this is a somewhat subjective art, with a goal of
% not allowing the GP to become so flexible that it does not require
% periodicity to fit the data even in the presence of clear periodicity.
In particular, we find it necessary to avoid large values of
$A$ and $\Gamma$, and small values of $l$; though the exact details of what
works well may differ among datasets (see \ref{sec:} and
\ref{sec:} for two different examples of hyperparameter priors).
More physically motivated priors are discussed in \textsection
\ref{sec:discussion}.

@@

We do not attempt to decide which single
peak in the ACF best represents the true rotation period,
but rather we identify several \emph{candidate} periods and define
a weighting scheme in order to create a noncommital, though useful,
multimodal prior.  While this procedure does not avoid heuristic choices,
we soften the potential impact of these choices because we are simply
creating a \emph{prior} for probabilistically justified
inference rather attempting to identify a single correct period.
